Fast and Error-Adaptive Influence Maximization based on Count-Distinct Sketches

### 2022-10-13 paper.tex
progress

\maketitle
\begin{abstract}
%     \fixme{
% Influence maximization~(IM) is the problem of finding a seed vertex set that maximizes the expected number of vertices influenced under a given diffusion model. Due to the NP-Hardness of finding an optimal seed set, approximation algorithms are frequently used for IM. }
% In addition to these high-quality yet expensive approximation algorithms, lightweight, sketch-based approaches which do not step-by-step simulate the diffusion process have been proposed in the literature to cope with the scale of today's networks. 
In this work, we describe a fast, error-adaptive approach that leverages Count-Distinct sketches and hash-based fused sampling to estimate the number of influenced vertices throughout a diffusion. We use per-vertex Flajolet-Martin sketches where each sketch corresponds to a sampled subgraph. To efficiently simulate the diffusions, the reach-set cardinalities of a single vertex are stored in memory in a consecutive fashion. This allows the proposed algorithm to estimate the number of influenced vertices in a single step for simulations at once. %In addition, thanks to their efficiency, and the scalability of our parallel implementation, we can rebuild the sketches whenever it is necessary. 
For a faster IM kernel, we rebuild the sketches in parallel only after observing estimation errors above a given threshold. Our experimental results show that the proposed algorithm yields comparable seed sets while being up to $3,337\times$ faster than a state-of-the-art, high-quality influence maximization algorithm. In addition, it is up to $63\times$ faster than a sketch-based approach while producing seed sets with $2\%$--$10\%$ better influence scores.
\end{abstract}
% } 
% \begin{keyword}
% Influence Maximization, Graph Processing, Graph Sampling, Fused Sampling, Memory Access Regularization, Count-Distinct Sketch
% \end{keyword}
% \end{frontmatter}

@@

Throughout the process, each register is used only for a single sample/simulation. 
For an edge ($u, v$) in simulation $j$ where $v$ is a live vertex, an update, i.e., a merge operation, on $u$'s register is performed on the corresponding register $M_u$[$j$], i.e., $M_u[j] = max(M_u[j],M_v[j])$. 
The toy example in Fig.~\ref{fig:fmops} shows how the sketch operations are done on a toy problem. In Fig.~\ref{fig:fmops}(a), the sketches, in a toy problem with 4 samples and 4-bit hash functions, are initialized; the id of vertex $u$ is hashed and leading zeros of the hash are written to registers. In Fig.~\ref{fig:fmops}(a), the merge operation is illustrated on registers of vertex $u$, $M_u$, and vertex $v$, $M_v$; merge operation takes element-wise maximum of registers to find union of the reachability sets. Fig.~\ref{fig:fmops} shows writing the merged reachability sets back to registers w.r.t sampling using a conditional move operation.

@@

 Algorithm~\ref{algo:main} first initializes the reachability sets of all vertices by adding the vertices themselves. 
 That is for all vertices $u$, its $j$th register is set to $M_u[j]=clz(h_j(u))$ meaning $R_{{G}_j}(u) = \{u\}$ where $G_j$ is the $j$th sampled graph. 
 Then, we perform the diffusion process on the sketch registers whose pseudocode is given in Algorithm~\ref{algo:diffusion-step}. The diffusion starts by adding all the vertices to the {\em live vertex set} $L$. 
 Then at each step, the incoming edges of the live vertices are processed. 
 For a vertex $u$, its sketch, $M_u$, is updated by merging the sketches $M_v$ of all live outgoing neighbors vertices $v \in {L} \cap \Gamma^+_{G}(u)$. For each such vertex $v$ and simulation $j$, the operation $M_u[j] = max(M_u[j], M_v[j])$ is performed. This approach can be seen as a bottom-up, i.e., reversed,  diffusion process where at each iteration, the cardinality information is pulled from vertices neighbors.
 \begin{figure}[!ht]
    \begin{center}
    \includegraphics[width=\linewidth]{images/infuserdiff.pdf}
    \caption{InFuseR diffusion process illustrated for only vertex $1$ on toy sample graph.  (a) \label{fig:infuserinit} Toy graph InFuseR initialization.  (b) \label{infuserdiff1} First level of diffusion using \emph{push-based} method for vertex $1$. (c) Second level of diffusion following Fig.\ref{fig:infuserdiff}(a).
     }\label{fig:infuserdiff} 
    \end{center}
\end{figure}
 \begin{figure}[!ht]
    \begin{center}
    \includegraphics[width=\linewidth]{images/sketchdiff.pdf}
    \caption{\acro diffusion process illustrated for \emph{all} vertices on toy sample graph. (a)\label{hyperinit} Toy sketch initialization. (b) \label{hyperdiff1} First level of diffusion using \emph{Pull-based} method for \emph{all} vertices. (c) \label{hyperdiff2} Second level diffusion following Fig.\ref{fig:hyperdiff}(a), only updated vertices' sketches are pulled. 
     }\label{fig:hyperdiff} 
    \end{center}
\end{figure}

@@

The traditional Greedy algorithm~\cite{kempe2003maximizing} processes the simulations one-by-one and computes the vertices' reachability sets. On the other hand, \acro efficiently performs multiple simulations at once in a single-step iteration (which processes distance-1 neighborhoods). Since each iteration relays one level of cardinality information, this step requires at most $d$ iterations where $d$ is the diameter of $G$. Note that when processed individually (as the Greedy algorithm), the $j$th simulation over the sampled subgraph $G_j$ would require only at most $d_j$ iterations, which is the diameter of $G_j$, and probably much smaller than $d$ considering the sampled subgraphs being highly sparse. However, $d$ is also a loose upper bound for \acro. A better one is {\tt max}$\{d_j: 1 \leq j \leq {\cal J}\} + 1$ where $d_j$ is the diameter of $G_j$. We have observed that during these simulations later iterations have only a minor impact on the marginal influence scores. Hence, to further reduce the overhead of concurrent simulation processing and avoid bottleneck simulations due to remaining perimeter structures (e.g., paths on the perimeter), we employ an early-exit threshold $\epsilon_c$ over the remaining live vertices ratio, which is expected to be very small when only one or two (bottleneck) simulations remain. That is if $|L'| \leq |V| \times \epsilon_c$ the diffusion process in Algorithm~\ref{algo:diffusion-step} stops. Otherwise, $L$ is set to $L'$, $L'$ is cleared, and the next iteration starts. We used $\epsilon_c = 0.02$ to make \acro faster while keeping its quality almost the same. Figure~\ref{fig:hyperdiff} shows a toy problem diffusion process on 4 vertices and a single sample, the process initializes sketches using respective hashes' leading zeros and bigger numbers are pulled from all recently changed neighbors. Whereas Infuser works vertices one by one, labeling visited vertices for the processed vertex. The process is repeated for multiple vertices until a seed vertex is found. Figure~\ref{fig:infuserdiff} shows an example in the same settings as Figure~\ref{fig:hyperdiff} to emphasize the difference between the two algorithms.

>>>>

\maketitle
\begin{abstract}
%     \fixme{
Influence maximization~(IM) is the problem of finding a seed vertex set that maximizes the expected number of vertices influenced under a given diffusion model. Due to the NP-Hardness of finding an optimal seed set, approximation algorithms are frequently used for IM. 
% }
In addition to these high-quality yet expensive approximation algorithms, lightweight, sketch-based approaches which do not step-by-step simulate the diffusion process have been proposed in the literature to cope with the scale of today's networks. 
In this work, we describe a fast, error-adaptive approach that leverages Count-Distinct sketches and hash-based fused sampling to estimate the number of influenced vertices throughout a diffusion. We use per-vertex Flajolet-Martin sketches where each sketch corresponds to a sampled subgraph. To efficiently simulate the diffusions, the reach-set cardinalities of a single vertex are stored in memory in a consecutive fashion. This allows the proposed algorithm to estimate the number of influenced vertices in a single step for simulations at once. %In addition, thanks to their efficiency, and the scalability of our parallel implementation, we can rebuild the sketches whenever it is necessary. 
For a faster IM kernel, we rebuild the sketches in parallel only after observing estimation errors above a given threshold. Our experimental results show that the proposed algorithm yields comparable seed sets while being up to $3,337\times$ faster than a state-of-the-art, high-quality influence maximization algorithm. In addition, it is up to $63\times$ faster than a sketch-based approach while producing seed sets with $2\%$--$10\%$ better influence scores.
\end{abstract}
% } 
% \begin{keyword}
% Influence Maximization, Graph Processing, Graph Sampling, Fused Sampling, Memory Access Regularization, Count-Distinct Sketch
% \end{keyword}
% \end{frontmatter}

@@

Throughout the process, each register is used only for a single sample/simulation. 
For an edge ($u, v$) in simulation $j$ where $v$ is a live vertex, an update, i.e., a merge operation, on $u$'s register is performed on the corresponding register $M_u$[$j$], i.e., $M_u[j] = max(M_u[j],M_v[j])$. 
The toy example in Fig.~\ref{fig:fmops} shows how the sketch operations are done on a toy problem. In Fig.~\ref{fig:fmops}(a), the sketches, in a toy problem with 4 samples and 4-bit hash functions, are initialized; the id of vertex $u$ is hashed and leading zeros of the hash are written to registers. In Fig.~\ref{fig:fmops}(a), the merge operation is illustrated on registers of vertex $u$, $M_u$, and vertex $v$, $M_v$; merge operation takes element-wise maximum of registers to find union of the reachability sets. Fig.~\ref{fig:fmops} shows writing the merged reachability sets back to registers w.r.t sampling using a conditional move operation. In the example, we see four possible situations that can happen; The operations on the first registers show moving bigger register $M_v[1]$ to $M_u[0]$. The second one shows the opposite, a bigger number is on the target register and there is no edge between $u$ and $v$ in sample graph $\hat{G}$. The third one shows an example when $M_v[3]$ is bigger than $M_u[3]$ but there is no edge between $u$ and $v$ in sample graph $\hat{G}$. The operations on the fourth register show when there is an edge between $u$ and $v$ but $M_u[4]$ is smaller than $M_v[4]$ so the move operation is not needed.

@@

 Algorithm~\ref{algo:main} first initializes the reachability sets of all vertices by adding the vertices themselves. 
 That is for all vertices $u$, its $j$th register is set to $M_u[j]=clz(h_j(u))$ meaning $R_{{G}_j}(u) = \{u\}$ where $G_j$ is the $j$th sampled graph. 
 Then, we perform the diffusion process on the sketch registers whose pseudocode is given in Algorithm~\ref{algo:diffusion-step}. The diffusion starts by adding all the vertices to the {\em live vertex set} $L$. 
 Then at each step, the incoming edges of the live vertices are processed. 
 For a vertex $u$, its sketch, $M_u$, is updated by merging the sketches $M_v$ of all live outgoing neighbors vertices $v \in {L} \cap \Gamma^+_{G}(u)$. For each such vertex $v$ and simulation $j$, the operation $M_u[j] = max(M_u[j], M_v[j])$ is performed. This approach can be seen as a bottom-up, i.e., reversed,  diffusion process where at each iteration, the cardinality information is pulled from vertices neighbors.

@@

The traditional Greedy algorithm~\cite{kempe2003maximizing} processes the simulations one-by-one and computes the vertices' reachability sets. On the other hand, \acro efficiently performs multiple simulations at once in a single-step iteration (which processes distance-1 neighborhoods). Since each iteration relays one level of cardinality information, this step requires at most $d$ iterations where $d$ is the diameter of $G$. Note that when processed individually (as the Greedy algorithm), the $j$th simulation over the sampled subgraph $G_j$ would require only at most $d_j$ iterations, which is the diameter of $G_j$, and probably much smaller than $d$ considering the sampled subgraphs being highly sparse. However, $d$ is also a loose upper bound for \acro. A better one is {\tt max}$\{d_j: 1 \leq j \leq {\cal J}\} + 1$ where $d_j$ is the diameter of $G_j$. We have observed that during these simulations later iterations have only a minor impact on the marginal influence scores. Hence, to further reduce the overhead of concurrent simulation processing and avoid bottleneck simulations due to remaining perimeter structures (e.g., paths on the perimeter), we employ an early-exit threshold $\epsilon_c$ over the remaining live vertices ratio, which is expected to be very small when only one or two (bottleneck) simulations remain. That is if $|L'| \leq |V| \times \epsilon_c$ the diffusion process in Algorithm~\ref{algo:diffusion-step} stops. Otherwise, $L$ is set to $L'$, $L'$ is cleared, and the next iteration starts. We used $\epsilon_c = 0.02$ to make \acro faster while keeping its quality almost the same. Figure~\ref{fig:hyperdiff} shows a toy problem diffusion process on 4 vertices and a single sample, the process initializes sketches using respective hashes' leading zeros and bigger numbers are pulled from all recently changed neighbors. Whereas Infuser works vertices one by one, labeling visited vertices for the processed vertex. The process is repeated for multiple vertices until a seed vertex is found. Figure~\ref{fig:infuserdiff} shows an example in the same settings as Figure~\ref{fig:hyperdiff} to emphasize the difference between the two algorithms. Please notice that in the given examples, \acro{} computes estimated influence for \emph{all} vertices, whereas InFuseR only computes influence for vertex $1$. Due to this reason, on Figure~\ref{fig:infuserdiff} reachability sets are shown for all vertices with dense vectors, the sets that are not currently manipulated are greyed out in the figure.
\begin{figure}[!ht]
    \begin{center}
    \includegraphics[width=\linewidth]{images/sketchdiff.pdf}
    \caption{\acro diffusion process illustrated for \emph{all} vertices on toy sample graph. (a)\label{hyperinit} Toy sketch initialization. (b) \label{hyperdiff1} First level of diffusion using \emph{Pull-based} method for \emph{all} vertices. (c) \label{hyperdiff2} Second level diffusion following Fig.\ref{fig:hyperdiff}(a), only updated vertices' sketches are pulled. 
     }\label{fig:hyperdiff} 
    \end{center}
\end{figure}
\begin{figure}[!ht]
    \begin{center}
    \includegraphics[width=\linewidth]{images/infuserdiff.pdf}
    \caption{InFuseR diffusion process illustrated for \emph{only vertex $1$} on toy sample graph.  (a) \label{fig:infuserinit} Toy graph InFuseR initialization.  (b) \label{infuserdiff1} First level of diffusion using \emph{push-based} method for vertex $1$. (c) Second level of diffusion following Fig.\ref{fig:infuserdiff}(a).
     }\label{fig:infuserdiff} 
    \end{center}
\end{figure}
